---
title: "Evo_Geno_Lab5x"
author: "Danaiijah Vilsaint"
date: "9/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(tidyverse)
```
```{r}
library(lubridate)
```
## Confirmed
```{r,eval=FALSE}
time_series_confirmed_long<- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")  %>% 
               pivot_longer(-c(Province_State, Country_Region, Lat, Long),
                             names_to = "Date", values_to = "Confirmed")
```
### Deaths
```{r,eval=FALSE}
time_series_deaths_long <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")  %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long),
               names_to = "Date", values_to = "Deaths")
```

##Key
```{r,eval=FALSE}
time_series_confirmed_long <- time_series_confirmed_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".", remove = FALSE)
time_series_deaths_long <- time_series_deaths_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".") %>% 
  select(Key, Deaths)
```
##Join tables
```{r,eval=FALSE}
time_series_long_joined <- full_join(time_series_confirmed_long,
    time_series_deaths_long, by = c("Key")) %>% 
    select(-Key)
```
##Reformat the data
```{r,eval=FALSE}
time_series_long_joined$Date <- mdy(time_series_long_joined$Date)
```

## Create Report table with counts not working
```{r, code=FALSE}
pivot_longer(-c( Province_State, Country_Region, Lat, Long, Date),
               names_to = "Report_Type", values_to = "Counts")
```
```{r}
library(gganimate)
```
```{r}
library(transformr)
```
```{r}
theme_set(theme_bw())
```
```{r}
library(gifski)
```



```{r, code=FALSE}
data_time <- time_series_long_joined %>% 
    group_by(Country_Region,Date) %>% 
    summarise_at(c("Confirmed", "Deaths"), sum) %>% 
    filter (Country_Region %in% c("China","Korea, South","Japan","Italy","US")) 
p<- ggplot(data_time, aes(x = Date,  y = Confirmed, color = Country_Region)) + 
      geom_point() +
      geom_line() +
      ggtitle("Confirmed COVID-19 Cases") +
      geom_point(aes(group = seq_along(Date))) +
  transition_reveal(Date)
animate(p,renderer = gifski_renderer(),end_pause = 15) 
```
```{r,code=FALSE}
data_time <- time_series_long_joined %>% 
    group_by(Country_Region,Date) %>% 
    summarise_at(c("Confirmed", "Deaths"), sum) %>% 
    filter (Country_Region %in% c("China","Korea, South","Japan","Italy","US")) 
p <- ggplot(data_time, aes(x = Date,  y = Confirmed, color = Country_Region)) + 
      geom_point() +
      geom_line() +
      ggtitle("Confirmed COVID-19 Cases") +
      geom_point(aes(group = seq_along(Date))) +
      transition_reveal(Date)
animate(p,renderer = gifski_renderer(),end_pause = 15) 
```
## Challenge 1:Print a graph (different from the one above) to a png file using 3*ppi for the height and width and display the png file in the report using the above R Markdown format.
```{r, exercise1, fig.width = 3, fig.height = 3,code=FALSE}
data_time <- time_series_long_joined %>% 
    group_by(Country_Region,Date) %>% 
    summarise_at(c("Confirmed", "Deaths"), sum) %>% 
    filter (Country_Region %in% c("China","Korea, South","Japan","Italy","US")) 
p <- ggplot(data_time, aes(x = Date,  y = Confirmed, color = Country_Region)) + 
      geom_point() +
      geom_line() +
      ggtitle("Confirmed COVID-19 Cases") +
      geom_point(aes(group = seq_along(Date))) +
      transition_reveal(Date)
animate(p,renderer = gifski_renderer(),end_pause = 15) 
```
## Challenge 2 : Turn one of the exercises from Lab 5 into an interactive graph with plotyly
```{r, code=FALSE}
Confirmed <- time_series_long_joined %>% arrange(desc(Confirmed)) %>% slice(1:10)
time_series_long_joined %>% 
    group_by(Country_Region,Date) %>% 
    summarise_at(c("Confirmed", "Deaths"), sum) %>% 
    filter (Country_Region %in% c("Brazil","UK","Italy", 
                                "Mexico", "US","Spain","France","Russia","Germany","China")) 
   p<- ggplot(data_time,aes(x = Date,  y = Deaths, color = Country_Region)) + 
    geom_point() +
    geom_line() +
    ggtitle("COVID-19 Deaths")
geom_point(aes(group = seq_along(Date))) +
      transition_reveal(Date)
animate(p,renderer = gifski_renderer(),end_pause = 15) 
```

## Challenge 3: Create an animated graph of your choosing using the time series data to display an aspect (e.g. states or countries) of the data that is important to you.
```{r,code=FALSE}
data_time <- time_series_long_joined %>% 
    group_by(Country_Region,Date) %>% 
    summarise_at(c("Confirmed", "Deaths"), sum) %>% 
    filter (Country_Region %in% c("China","US")) 
p<- ggplot(data_time, aes(x = Date,  y = Confirmed, color = Country_Region)) + 
      geom_point() +
      geom_line() +
      ggtitle("Confirmed COVID-19 Cases in China and the US") +
      geom_point(aes(group = seq_along(Date))) +
  transition_reveal(Date)
animate(p,renderer = gifski_renderer(),end_pause = 15) 
```
