---
title: "Evo_Geno_Lab9"
author: "Danaiijah Vilsaint"
date: "10/26/2020"
output: html_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(stringr)
library(tidyverse)
```

### 14.3.1 Basic matches
```{r, exact strings}
x <- c("apple", "banana", "pear")
str_view(x, "an")
str_view(x, ".a.")
```


##14.3.1.1 Exercises 1
```{r}

#They dont match because there is not of the slashes you need for of them represent one
```


##14.3.2 Anchors
```{r}
x <- c("apple", "banana", "pear")
str_view(x, "^a")
```
```{r}
str_view(x, "e$")
```
##EX1
```{r}
#You match the literal string "$^$"?
```
##14.3.3.1 Character classes and alternatives
##Create regular expressions to find all words that:
Start with a vowel.
That only contain consonants. (Hint: thinking about matching “not”-vowels.)
End with ed, but not with eed.
End with ing or ise.
```{r}
x <- c("bed", "reading", "ng","apple","arise")
str_view(x, "ing$")
```
```{r}
str_view(x, "ing$")
```
```{r}
str_view(x, "[aeiou]")
```


14.3.4 Repetition
## Describe the equivalents of ?, +, * in {m,n} form.
```{r}
# ?: {0,1}
str_view(x, "C{0,1}")
str_view(x, 'C{0,1}?')
# +: {1,}
str_view(x, "C{1,}")
str_view(x, "C+")
# *: {0,}
str_view(x, "C{0,}")
str_view(x, "C*")
```
```

14.3.5 Grouping and backreferences
## Describe, in words, what these expressions will match:
(.)\1\1
"(.)(.)\\2\\1"
(..)\1
"(.).\\1.\\1"
"(.)(.)(.).*\\3\\2\\1"

```{r}
## They match words that have repetition within it
```

14.4.1 Detect matches
## For each of the following challenges, try solving it by using both a single regular expression, and a combination of multiple str_detect() calls.
Find all words that start or end with x.
Find all words that start with a vowel and end with a consonant.
Are there any words that contain at least one of each different vowel?
```{r}
x <- c(" What can I say except Youre welcome For the tides the sun the sky Hey it is okay  it is okay Youre welcome Im just an ordinary demi guy")
# Replace all forward slashes in a string with backslashes.
```

14.4.2 Extract matches
## In the previous example, you might have noticed that the regular expression matched “flickered”, which is not a colour. Modify the regex to fix the problem.
```{r}

```

14.4.3 Grouped matches
```{r}
# Find all words that come after a “number” like “one”, “two”, “three” etc. Pull out both the number and the word.
numbers1 <- "\\b(one|two|three|four|five|six|seven|eight|nine|ten) +(\\w+)"
sentences[str_detect(sentences, numbers1)] %>%
  str_extract(numbers1)
```


14.4.5 Splitting
```{r}
```{r, exercise10}
# Split up a string like “apples, pears, and bananas” into individual components.
y <- "apples, pears, and bananas."
str_view_all(y, boundary("word"))
```
```



