---
title: "Evo_Geno_Lab9x"
author: "Danaiijah Vilsaint"
date: "10/28/2020"
output: html_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
library(knitr)
```



```{r}
x=1
if (x == 1) { y = 3
       
} else {y = 2}
y
```
## example of a valid if/else structure
```{r}
## Generate a uniform random number
x <- runif(1, 0, 10)  
if(x > 3) {
        y <- 10
} else {
        y <- 0
}
```



##Loops
```{r}
for(i in 1:13) {
        print(i)
}
```
## Print out each element of 'x'
```{r}
x <- c("Frozen", "Moana", "Coco", "Tangled")
for(i in 1:4) {
        print(x[i])  
}

```
## Generate a sequence based on length of 'x'
```{r}
for(i in seq_along(x)) {   
        print(x[i])
}
```

##
```{r}
for(letter in x) {
        print(letter)
}
```
##For one line loops, the curly braces are not strictly necessary.
```{r}
for(i in 1:4) print(x[i])
```

## Nested for loops
```{r}
x <- matrix(1:6, 2, 3)
for(i in seq_len(nrow(x))) {
        for(j in seq_len(ncol(x))) {
                print(x[i, j])
        }   
}
```
## while Loops
```{r}
count <- 0
while(count < 10) {
        print(count)
        count <- count + 1
}
```
##Sometimes there will be more than one condition in the test.
```{r}
z <- 5
set.seed(1)
while(z >= 3 && z <= 10) {
        coin <- rbinom(1, 1, 0.5)
        
        if(coin == 1) {  ## random walk
                z <- z + 1
        } else {
                z <- z - 1
        } 
}
print(z)
```


## non-trivial function body
```{r}
f <- function() 
{cat("How Far I'll Go!\n")}
f()
```
## function arguments:determines how many times “Hello, world!” is printed to the console
```{r}
f <- function(num)
{for(i in seq_len(num))
cat("Into The Unknown!\n")}
f(3)
```
## This next function returns the total number of characters printed to the console
```{r}
f <- function(num = 4) 
{ Poco <- "Un Poco Loco!\n"
for(i in seq_len(num)) 
{cat(Poco)}
chars <- nchar(Poco) * num 
chars}
 meaningoflife <- f(3)
print(meaningoflife)
```
## named arguments 
```{r}
f(num = 2)
```


